---
title: "Synapse Tables - Changing a Column"
author: "Kenneth Daily"
date: "01/23/2015"
output: html_document
---

An example of changing a column type, keeping the same data.

First, set the column name that you want to change, and the table it comes from.

```{r}
columnToChange <- "Replicate"
tableId <- "syn3131513"
```

Get the existing data.

```{r}
## Get the existing data
oldTable <- synGet(tableId)
oldDataQuery <- synTableQuery(sprintf("SELECT * FROM %s", oldTable$properties$id))
oldData <- oldDataQuery@values
```

Now that you have the data, figure out how you delete a column. Currently, I do it in the web client.

After deleting the column, get the table again:

```{r}
newTable <- synGet(tableId)
```

Make a new column with same name and save it to Synapse. In this case I'm just changing the maximum size.

```{r}
newrepcol <- TableColumn(name=columnToChange, columnType="STRING", maximumSize=50, defaultValue="")
newcol <- synStore(newrepcol)
```

Add that column to the table and save:

```{r}

newTable$properties$columnIds <- c(newTable$properties$columnIds, newcol@id)
newTable <- synStore(newTable)

```

Now, make changes and save to Synapse. In this case I just copy the old data to the new column.

```{r}
newQuery <- synTableQuery(sprintf("SELECT * FROM %s", sampleProcessMetadataTable$properties$id))
newQuery@values[columnToChange] <- oldData$Replicate

## Commit the changes
synStore(newQuery)
```
